---
title: "Simulation"
output: html_notebook
---

```{r}
rm(list=ls())

library(tidyverse)
library(MASS)

# functions
constrain <- function(v, minV, maxV) {
  sapply(v, function(y) min(max(y, minV), maxV))
}

# load data
load(file="data/analytic.RData")
```

# Basics of Nested Simulation

### Generating the identifiers

```{r}
# calculate N & M
N <- sum(analytic$enrollment_cnt)
M <- length(analytic$enrollment_cnt)

# generate student ids
stid <- 1:N

# generate school ids for students
schid <- rep(analytic$schid, analytic$enrollment_cnt)

# test
df <- data.frame(stid, schid)
```

### Simulating from aggregate parameters

#### Continuous

```{r}
attend_rate_1 <- rnorm(analytic$enrollment_cnt[1], analytic$attend_rate_pct[1], 15)

summary(attend_rate_1)

attend_rate_1 <- constrain(attend_rate_1, 0, 100)

summary(attend_rate_1)
```

#### Discrete

```{r}
farms_1 <- as.integer(runif(analytic$enrollment_cnt[1]) <= analytic$farms_pct[1]/100)

summary(farms_1)

farms_1 <- rbinom(analytic$enrollment_cnt[1], 1, analytic$farms_pct[1]/100)

summary(farms_1)
```


### Simulating from vectors of aggregate parameters

```{r}
sch_attend_rate <- rep(analytic$attend_rate_pct, analytic$enrollment_cnt)

attend_rate <- sch_attend_rate %>%
  rnorm(sch_attend_rate, 15) %>% 
  constrain(0, 100)

summary(attend_rate)

sch_farms_pct <- rep(analytic$farms_pct, analytic$enrollment_cnt)

farms <- rbinom(sch_farms_pct, 1, sch_farms_pct/100)

summary(farms)

# test
df <- data.frame(stid, schid, attend_rate, farms)
```

#### Examine aggregates

```{r}
View(
  left_join(
    dplyr::select(analytic, schid, schname, attend_rate_pct, farms_pct),
    df %>% group_by(schid) %>% summarize(
      attend_rate_pct_sim = mean(attend_rate),
      farms_pct_sim = mean(farms)*100
    ),
    by="schid"
  ) %>% dplyr::select(schid, schname, attend_rate_pct, attend_rate_pct_sim, farms_pct, farms_pct_sim)
)
```

#### Examine correlation

```{r}
with(df, cor(attend_rate, farms))
```

# Parameter Distributions {.tabset}

```{r}
plot_dist <- function(v, discrete=F) {
  if (discrete) {
    data.frame(x = v) %>%
      ggplot(aes(x = x)) +
        geom_bar() +
        theme_minimal()
  } else {
    data.frame(x = v) %>%
      ggplot(aes(x = x)) +
        geom_line(stat="density") +
        theme_minimal()
  }
}
```


## Uniform

```{r, results="asis"}
plot_dist(runif(100000, 0, 1))
```

## Normal

```{r, results="asis"}
plot_dist(rnorm(100000, 100, 25))
```

## Bernoulli

```{r}
plot_dist(rbinom(100000, 1, 0.7), discrete=T)
```

## Binomial

```{r}
plot_dist(rbinom(100000, 10, 0.5), discrete=T)
```

## Beta

```{r}
plot_dist(rbeta(100000, 2, 6))
```

## Poison

```{r}
plot_dist(rpois(100000, 3), discrete=T)
```

## Gamma

```{r}
plot_dist(rgamma(100000, 5, 2))
```

## attend_rate

```{r}
plot_dist(df$attend_rate)
```

## farms

```{r}
plot_dist(df$farms, discrete = T)
```

# Simulating Covariation

### Simulating from aggregate parameters

```{r}
r <- 0.6*sqrt(15)*sqrt(10)

df <- mvrnorm(analytic$enrollment_cnt[1], 
              mu = c(analytic$attend_rate_pct[1], 0),
              Sigma=matrix(c(15, r, r, 10), 2, 2)) %>% data.frame

colnames(df) <- c("attend_rate", "relative_income")

df$stid <- 1:analytic$enrollment_cnt[1]

df <- arrange(df, relative_income)

df$farms <- rep(c(1, 0), 
                c(
                  round(analytic$enrollment_cnt[1]*analytic$farms_pct[1]/100),
                  analytic$enrollment_cnt[1] - round(analytic$enrollment_cnt[1]*analytic$farms_pct[1]/100)
                  )
)

df <- df %>% dplyr::select(stid, attend_rate, farms) %>% arrange(stid)
```

### Simulating from vectors of aggregate parameters

More complicated!

