---
title: "Baltimore High Schools Analysis"
author: "Nathaniel A. Dewey"
output:
  html_notebook:
    code_folding: hide
---

```{r}
rm(list=ls())

# install.packages("devtools")
# devtools::install_github("thomasp85/patchwork")

library(tidyverse)
library(summarytools)
library(stargazer)
library(patchwork)

# load data
load(file="data/analytic.RData")

analytic <- analytic %>% mutate_at(vars(-schid, -schname, -schtype, -star_rating, -points_earned), function(x) {
  x %>%
    ifelse(
      . == "<= 5.0",
      round(runif(nrow(analytic), 0, 5), 1),
      .
      ) %>%
    ifelse(
      . == ">= 95.0",
      round(runif(nrow(analytic), 95, 100), 1),
      .
      ) %>%
    as.numeric
})
```

## Analytics

### Descriptive Statistics

```{r}
with(analytic, descr(points_earned))
```

### Frequency Table

```{r}
with(analytic, freq(star_rating))
```


### Contingency Table

```{r}
with(analytic, ctable(star_rating, schtype, prop="c"))
```

### Correlation Table

```{r}
with(analytic, cor(select(analytic, farms_pct, chronic_absentee_pct, mcap_algebra_proficient_pct, mcap_ela10_proficient_pct)))
```

### Linear Regression

```{r}
fit_lr <- lm(points_earned ~ farms_pct, data=analytic)

summary(fit_lr)
```

### Multiple Regression

```{r}
fit_mr <- lm(farms_pct ~ mcap_algebra_proficient_pct + mcap_ela10_proficient_pct, data=analytic)

summary(fit_mr)
```

### Regression Tables

```{r}
fit_model_1 <- lm(farms_pct ~ chronic_absentee_pct + mcap_algebra_proficient_pct, data=analytic)
fit_model_2 <- lm(farms_pct ~ chronic_absentee_pct + mcap_ela10_proficient_pct, data=analytic)
fit_model_3 <- lm(farms_pct ~ chronic_absentee_pct + mcap_algebra_proficient_pct + mcap_ela10_proficient_pct, data=analytic)

suppressWarnings(
  stargazer(fit_model_1, fit_model_2, fit_model_3, type="text")
)
```

## Visualization

### Univariate Plots {.tabset}

#### Histogram

```{r}
ggplot(analytic) +
  geom_histogram(aes(x=mcap_ela10_proficient_pct), bins=20)
```

#### Density

```{r}
ggplot(analytic) +
  geom_density(aes(x=mcap_ela10_proficient_pct))
```

#### Better Density

```{r}
analytic %>%
  pivot_longer(c(mcap_ela10_proficient_pct, mcap_algebra_proficient_pct), names_to="mcap_test", values_to="mcap_proficient_pct") %>%
  ggplot +
    geom_line(aes(x=mcap_proficient_pct, color=mcap_test), stat="density")
```

#### Box Plots

```{r}
analytic %>%
  pivot_longer(c(mcap_ela10_proficient_pct, mcap_algebra_proficient_pct), names_to="mcap_test", values_to="mcap_proficient_pct") %>%
  ggplot +
    geom_boxplot(aes(x=mcap_test, y=mcap_proficient_pct))
```

### Multivariate Plots {.tabset}

#### Scatter Plot

```{r}
ggplot(analytic) +
  geom_point(aes(y = (100 - mcap_ela10_proficient_pct), x = farms_pct))
```

```{r}
ggplot(analytic) +
  geom_point(aes(y = (100 - mcap_ela10_proficient_pct), x = farms_pct)) +
  geom_point(aes(y = (100 - mcap_ela10_proficient_pct), x = farms_pct), data=function(df) { filter(df, schid %in% c(403, 407, 415, 480)) }, color="red")
```

#### Scatter Plot with Jitter

```{r}
analytic %>%
  mutate(
    mcap_ela10_proficient_pct_cut = cut(mcap_ela10_proficient_pct, breaks=c(0, 25, 50, 75, 100), labels=F),
    farms_pct_cut = cut(farms_pct, breaks=c(0, 25, 50, 75, 100), labels=F)
  ) %>%
  ggplot +
    geom_jitter(aes(y=(100 - mcap_ela10_proficient_pct_cut), x=farms_pct_cut), width=0.1, height=0.1) +
    labs(y = "% Not Proficient on MCAP ELA", x = "% FARMS") +
    theme_minimal()
```

#### Scatter Plot with Linear Regression Fit

```{r}
ggplot(analytic, aes(y = (100 - mcap_ela10_proficient_pct), x = farms_pct)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE)
```

#### Scatter Plot with Polynomial Fit

```{r}
ggplot(analytic, aes(y = (100 - mcap_ela10_proficient_pct), x = farms_pct)) +
  geom_point() +
  geom_smooth(method = "lm", formula=y ~ poly(x, 2), se = TRUE)
```

#### Scatter Plot with Nonparametric Fit

```{r}
ggplot(analytic, aes(y = (100 - mcap_ela10_proficient_pct), x = farms_pct)) +
  geom_point() +
  geom_smooth(method = "loess", se = TRUE)
```

### Multiples

#### Faceting

```{r}
ggplot(analytic) +
  geom_point(aes(y = (100 - mcap_ela10_proficient_pct), x = farms_pct)) +
  facet_wrap(vars(schtype))
```


```{r}
ggplot(analytic) +
  geom_point(aes(y = (100 - mcap_ela10_proficient_pct), x = farms_pct)) +
  facet_grid(vars(schtype), vars(star_rating))
```

#### Patchwork

```{r}
p_H <- analytic %>%
  filter(schtype == "H") %>%
  ggplot +
    geom_point(aes(y = (100 - mcap_ela10_proficient_pct), x = farms_pct))

p_MH <- analytic %>%
  filter(schtype == "MH") %>%
  ggplot +
    geom_point(aes(y = (100 - mcap_ela10_proficient_pct), x = farms_pct))

p_H + p_MH
```


```{r}
p1 <- ggplot(analytic) +
  geom_point(aes(y = (100 - mcap_ela10_proficient_pct), x = farms_pct)) +
  scale_x_continuous(position = "top") +
  labs(x = "% Direct Certification", y = "% Below Proficient on MCAP ELA") +
  theme_minimal()

p2 <- ggplot(analytic) +
  geom_area(aes(x = (100 - mcap_ela10_proficient_pct)), stat="density") +
  coord_flip() +
  theme_void()

p3 <- ggplot(analytic) +
  geom_area(aes(x = farms_pct), stat="density") +
  scale_y_reverse() +
  theme_void()
  
(p1 | p2) / (p3 | plot_spacer())
```

```{r}
layout <- "
AAAAAB
AAAAAB
AAAAAB
AAAAAB
AAAAAB
CCCCC#
"

p1 + p2 + p3 +
  plot_layout(design = layout)
```

